%============================================================================
% LaTeX File
% Daniel J. Greenhoe 
%============================================================================

%=======================================
\chapter{Description}
%=======================================
\chardh{描}{寫}{mi\'ao}{x\v{i}e}{trace}{write}{description}

%=======================================
\section{Adjectives}
%=======================================
  \tcom{
    \chardh{形}{容}{x\'ing}{r\'ong}{appearance}{contain}{describe}
    \charsh{詞}{c\'i}{word}
    }{adjective}

\begin{tblvv}{Size}{voc:size}
   big             & \charsp{大}{d\`a    }{big}                                   &
   small           & \charsp{小}{x\v{i}ao}{small}                                 \tblh
   tall            & \charsp{高}{g\=ao   }{tall}                                  &
   short           & \charsp{矮}{a\v{i}  }{short}                                 \tblh
   fat             & \charsp{胖}{p\`ang}{fat}                                     &
   thin            & \charsp{瘦}{sh\`ou}{thin}                                    \tblh
   huge            & \chards{巨}{大}{j\`u}{d\`a}{huge}{big}{huge}                 &
   tiny            & \chards{微}{小}{w\'ei}{xi\v{a}o}{tiny}{small}{tiny}
\end{tblvv}
 
\begin{tblvv}{misc.}{voc:misc}
  \tblx beautiful       & \chards{美}{麗}{m\v{e}i}{l\`i}{beauty}{beautiful}{beautiful} 
  \tblc ugly            & \charsp{醜}{ch\v{o}u}{ugly}                                  
  \tblh pretty          & \chards{漂}{亮}{pi\`ao}{li\`ang}{pretty}{bright}{pretty}     
  \tblc handsome        & \charsp{帥}{sh\`uai}{handsome}                               
  \tblh smart           & \chards{聰}{明}{c\=ong}{mi\'ng}{hear well}{bright}{smart}    
  \tblc dumb            & \charsp{笨}{b\`en}{dumb}                                     
  \tblh sensible        & \chards{懂}{事}{d\v{o}ng}{sh\`i}{understand}{affairs}{sensible} 
  \tblh dress up        & \chards{打}{扮}{d\v{a}}{b\`an}{to hit}{to play the part of}{to dress up}
\end{tblvv}
 
\begin{tbls}{Pretty and handsome}{sen:pretty}
    You are very pretty.
    &
    \chars{你}{n\v{i}}{you} \chars{很}{h\v{e}n}{very} 
    \chard{漂}{亮}{pi\`ao}{li\`ang}{pretty}{bright}{pretty}.
  \tblh
    He is very handsome.
    &
    \chars {他}{t\=a}{he} \chars{很}{h\v{e}n}{very} 
    \chars{帥}{sh\`uai}{handsome}.
  \tblh
   She is green but very pretty.
   &
   \chars{她}{t\=a}{she} \chars{是}{sh\`i}{is} 
   \chard{綠}{色}{l\`v   }{s\`e}{green }{color}{green } \chars{的}{de}{'s}
   \chard{可}{是}{k\v{e}}{s\`hi}{approve}{that}{but}   
   \chars{很}{h\v{e}n}{very} 
   \chards{漂}{亮}{pi\`ao}{li\`ang}{pretty}{bright}{pretty}.
\end{tbls}


\chardh{顏}{色}{y\'an}{s\`e}{face}{color}{color}

\begin{tblvv}{Colors}{voc:colors}
   black          & \chards{黑}{色}{h\=ei  }{s\`e}{black }{color}{black } &
   white          & \chards{白}{色}{b\=ai  }{s\`e}{white }{color}{white } \tblh
   red            & \chards{紅}{色}{h\'ong }{s\`e}{red   }{color}{red   } &
   pink           & \chards{粉}{紅}{f\v{e}n}{h\'ong}{powder}{red}{pink} \chars{色}{s\`e}{color} \tblh
   green          & \chards{綠}{色}{l\`v   }{s\`e}{green }{color}{green } &
   blue           & \chards{藍}{色}{l\'an}{s\`e}{blue}{color}{blue}       \tblh
   yellow         & \chards{黃}{色}{hu\'ang}{s\`e}{yellow}{color}{yellow} &
   brown          & \chards{褐}{色}{h\'e   }{s\`e}{brown }{color}{brown } \tblh
   purple         & \chards{紫}{色}{z\v{i} }{s\`e}{purple}{color}{purple} &
   grey           & \chards{灰}{色}{h\=ui  }{s\`e}{grey  }{color}{grey  }
\end{tblvv} 


\begin{tbls}{Colors}{sen:colors}
    What color is it?
    &
    \chars{它}{t\=a}{it} \chars{是}{sh\`i}{is} 
    \chard{什}{麼}{sh\'e}{me}{what}{}{what}
    \chard{顏}{色}{y\'an}{s\`e}{face}{color}{color}?
  \tblh
   The ball is yellow and blue.
   & 
   \chars{球}{qi\'u}{ball} \chars{是}{sh\`i}{is} 
   \chards{黃}{色}{hu\'ang}{s\`e}{yellow}{color}{yellow}
   \chars{和}{h\`an}{and}
   \chards{黃}{色}{hu\'ang}{s\`e}{yellow}{color}{yellow}.
   \chars{的}{de}{'s}
\end{tbls}

Color Song  (Can be sung to the tune of {\em London Bridges})
\begin{tbls}{Color Song}{song:color}
    Red, yellow, green and blue;
    &
    \chars{紅}{h\'ong }{red}, \chars{黃}{hu\'ang}{yellow}, 
    \chars{綠}{l\`v}{green }, \chars{藍}{l\'an}{blue}
  \tblh
    green and blue; green and blue.
    &
    \chars{綠}{l\`v}{green }, \chars{藍}{l\'an}{blue}, 
    \chars{綠}{l\`v}{green }, \chars{藍}{l\'an}{blue}
  \tblh
    Red, yellow, green and blue;
    &
    \chars{紅}{h\'ong }{red}, \chars{黃}{hu\'ang}{yellow}, 
    \chars{綠}{l\`v}{green }, \chars{藍}{l\'an}{blue}
  \tblh
    White, black, purple, grey.
    &
    \chars{白}{b\=ai  }{white }, \chars{黑}{h\=ei  }{black },
    \chards{紫}{色}{z\v{i}}{s\`e}{purple}{color}{purple}, \chars{灰}{h\=ui  }{grey  }
\end{tbls}


%   \epsfig{file=../commonc/ukulele.eps,  width=2.5cm, clip= } \hfill
%   \epsfig{file=../commonc/biz_man.eps,  width=2.5cm, clip= } \hfill
%   \epsfig{file=../commonc/monster.eps,  width=2.5cm, clip= } \hfill
%   \epsfig{file=../commonc/bean.eps,     width=2.5cm, clip= } 
%\footnote{\url{http://www.animationfactory.com/}}




