%============================================================================
% LaTeX File
% Daniel Greenhoe
%============================================================================





%=======================================
\chapter{Simple conversation}
%=======================================


%=======================================
\section{Greetings}
%=======================================
\tcom{
  \zhtssH{打}{d\v{a}}{beat}
  \zhtsdH{招}{呼}{zh/-ao}{h/-u}{beckon}{exhale}{call}
  }{greeting}

\begin{tblss}{Hello and goodbye}{sen:grt_greetings}%
  \tblx hello             & \zhtss{你}{n\v{i}}{you}    \zhtss{好}{h\v{a}o}{good}%
  \tblc good bye          & \zhtsds{再}{見}{z\`ai}{ji\`an}{again}{see}{good bye}%
  \tblh good night        & \zhtsds{晚}{安}{w\v{a}n}{a\=n}{late}{peaceful}{good night}%
  \tblc see you tomorrow  & \zhtsd {明}{天}{mi\'ng}{ti\=an}{bright}{the sky}{tomorrow}\zhtss {見}{ji\`an}{to see}%
  \tblh see you later     & \zhtsd {一}{會兒}{y\=i}{h\v{u}i}{one}{to meet together}{in a little while}%
                            \zhtss{見}{ji\`an}{to see}%
  \tblc see you in a little while  & \zhtss{待}{d\`ai}{to wait for}%
                            \zhtss{會兒}{h\v{u}i}{to meet together}\zhtss{見}{ji\`an}{to see}%
  \tblh I'm out of here   & \zhtss{我}{w\v{o}}{I}\zhtss{走}{z\v{o}u}{to walk}\zhtssp{了}{le}{(indicates a change of state)}%
  \tblc                   &
\end{tblss}




\begin{tblss}{How are you?}{sen:grt_how_are_you}
  \tblx How are you?      & \zhtss{你}{n\v{i}}{you}    \zhtss{好}{h\v{a}o}{good}    \zhtss{嗎}{ma}{(?)}
  \tblc not bad           & \zhtss{不}{b\'u}{not}      \zhtss{錯}{c\`uo}{faults}
  \tblh I am fine.        & \zhtss{我}{w\v{o}}{I}      \zhtss{很}{h\v{e}n}{very}    \zhtss{好}{h\v{a}o}{good}
  \tblc so-so             & \zhtss{馬}{m\v{a}}{horse}  \zhtss{馬}{m\v{a}}{horse}
                            \zhtss{虎}{h\=u}{tiger}    \zhtss{虎}{h\=u}{tiger}
  \tblh thank you         & \zhtss{謝}{xi\`e}{to thank}\zhtss{謝}{xi\`e}{to thank}
  \tblc not too well      & \zhtss{不}{b\'u}{not}      \zhtss{太}{t\`ai}{extremely} \zhtss{好}{h\v{a}o}{good}
  \tblh And you?          & \zhtss{你}{n\v{i}}{you}    \zhtss{呢}{ne}{}
  \tblc                   &
\end{tblss}


%=======================================
\section{Introductions}
%=======================================
\begin{minipage}{\tw/4}
  \zhtsdH{介}{紹}{ji\`e}{sh\`ao}{situated between}{bring together}{introduce}
\end{minipage}

\begin{tbls}{Introductions}{sen:grt_intro}
  \tblx What is your name?
      & \zhtss{你}{n\v{i}}{you}%
        \zhtss{叫}{ji\`ao}{called}%
        \zhtsd{什}{麼}{sh\'e}{me}{what}{}{what}%
        \zhtsd{名}{字}{mi\'ng}{z\`i}{name}{a word}{a name}\zhtsQ%
  \tblh My name is \blank.
      & \zhtss{我}{w\v{o}}{I}\zhtss{叫}{ji\`ao}{called}\blank\zhtsP%
  \tblh What is his name?
      & \zhtss{他}{t\=a}{he}%
        \zhtss{叫}{ji\`ao}{called}%
        \zhtsd{什}{麼}{sh\'e}{me}{what}{}{what}%
        \zhtsd{名}{字}{mi\'ng}{z\`i}{name}{a word}{a name}\zhtsQ%
  \tblh His name is \blank.
      & \zhtss{他}{t\=a}{he}\zhtss{叫}{ji\`ao}{called}\blank\zhtsP%
  \tblh How old are you?
      & \zhtss{你}{n\v{i}}{you}\zhtss{幾}{j\v{i}}{how many}\zhtss{歲}{s\`ui}{a year of age}\zhtsQ%
  \tblh I am \blank years old.
      & \zhtss{我}{w\v{o}}{I}\blank\zhtss{歲}{s\`ui}{a year of age}\zhtsP%
  \tblh Can you speak English?
      & \zhtss{你}{n\v{i}}{you}\zhtss{會}{h\`ui}{able}\zhtss{說}{sh\=uo}{speak}%
        \zhtsd{英}{語}{yi\=ng}{y\v{u}}{outstanding}{language}{English}%
        \zhtss{嗎}{ma}{}\zhtsQ%
  \tblh It's nice to meet you.
      & \zhtss{很}{h\v{e}n}{very}\zhtsd{高}{興}{g\=ao}{xi\`ng}{high}{cheerful}{glad}%
        \zhtsd{認}{識}{r\`en}{sh\`i}{recognize}{know}{know}%
        \zhtss{你}{n\v{i}}{you}\zhtsP%
\end{tbls}






%=======================================
\section{Civilities}
%=======================================
\begin{minipage}{\tw/4}%
  \zhtsdH{禮}{貌}{l\v{i}}{m\`ao}{courtesy}{manner}{politeness}
\end{minipage}

\begin{tblvv}{Civilities}{voc:civilities}
  \tblx thanks  & \zhtsds{謝}{謝}{x\`ie}{x\`ie}{thank}{thank}{thanks}                         
  \tblc cordial & \zhtsds{親}{切}{qi\=n}{qi\`e}{intimate}{close}{cordial}                     
  \tblh manners & \zhtsds{禮}{貌}{l\v{i}}{m\`ao}{ceremony}{features}{manners}
\end{tblvv}


\begin{tbls}[3\tw/16]{Apologies}{sen:apology}
  \tblx sorry
      & \zhtss{對}{d\`ui}{facing} \zhtss{不}{b\`u}{not} \zhtss{起}{\v{q}i}{to rise}%
  \tblh sorry
      & \zhtsds{抱}{歉}{b\`ao}{qi\`an}{hug}{apology}{apologize}%
  \tblh no problem 
      & \zhtsd{沒}{有}{m\'ei}{y\v{o}u}{none}{have}{not have}%
        \zhtsd{問}{題}{w\`en}{t\'i}{to ask}{a subject}{problem}
  \tblh It doesn't matter.
      & \zhtss{沒}{m\'ei}{not}%
        \zhtsd{關}{係}{g\=uan}{x\`i}{to close}{to tie}{relation}\zhtsP
  \tblh Please don't be angry!%
      & \zhtss{請}{qi\v{n}g}{please}%
        \zhtsd{不}{要}{b\'u}{y\`ao}{not}{want}{don't}%
        \zhtsd{生}{氣}{sh\=eng}{q\`i}{to give birth to}{air}{angry}\zhtsX%
        \citetbl{\citerp{strange_visitor}{27}}%
  \tblh Don't take offense!%
      & \zhtss{別}{bi\'e}{don't}%
        \zhtsd{介}{意}{ji\`e}{y\`i}{introduce}{meaning}{to take offense}\zhtsX
  %\tblh
    %not have good intentions
    %&
    %\zhtss{懷}{h\'uai}{ming}
    %\zhtsd{好}{意}{h\v{a}o}{y\`i}{good}{meaning}{good intentions}
    %\citex{ugly_duckling}{24}
  \tblh I don't know what I should say to make things right.
      & \zhtss{我}{w\v{o}}{I}%
        \zhtss{不}{b\`u}{not}\zhtss{知}{zh\=i}{know}\zhtss{該}{g\=ai}{should}%
        \zhtss{說}{sh\=uo}{say}%
        \zhtsd{什}{麼}{sh\'e}{me}{what}{}{what}%
        \zhtss{才}{c\'ai}{just} \zhtss{好}{h\v{a}o}{good}\zhtsP%
        \citetbl{\citerp{butterfly_kite}{15}}%
  \tblh Together they were able to undo the misunderstanding.
      & \zhtsd{他}{們}{t\=a}{m\'en}{he}{s}{they}%
        \zhtsd{彼}{此}{b\v{i}}{\v{c}i}{that}{this}{each other}%
        \zhtsd{解}{開}{j\v{i}e}{k\=ai}{to untie}{to open}{to undo}%
        \zhtss{了}{le}{}%
        \zhtsd{誤}{會}{w\`u}{h\`ui}{error}{to assemble}{misunderstanding}\zhtsP%
        \citetbl{\citerp{3musketeers}{16}}%
  \tblh You must show good manners.
      & \zhtss{你}{\v{n}i}{you}%
        \zhtss{要}{y\`ao}{want}%
        \zhtsd{表}{現}{bi\v{a}o}{xi\`an}{surface}{to appear}{to show}%
        \zhtss{有}{y\v{o}u}{to have}%
        \zhtsds{禮}{貌}{l\v{i}}{m\`ao}{ceremony}{features}{manners}\zhtsP%
        \citetbl{\citerp{strange_visitor}{1}}%
\end{tbls}


\begin{tblss}{Civilities}{sen:civilities}
  \tblx thank you 
      & \zhtsds{謝}{謝}{x\`ie}{x\`ie}{thank}{thank}{thanks}\zhtss{你}{n\v{i}}{you}
  \tblc you're welcome 
      & \zhtss{不}{b\'u}{not}\zhtsd{客}{氣}{k\`e}{q\`i}{a guest}{air}{polite}%
  \tblh excuse me, can I ask \ldots%
      & \zhtss{請}{qi\v{n}g}{to request}\zhtss{問}{w\`en}{to request}\Large\ldots%
  \tblc excuse me (to pass)%
      & \zhtss{借}{ji\`e}{to borrow}\zhtss{過}{g\`uo}{to pass}%
  \tblh Welcome! 
      & \zhtsds{歡}{迎}{h\=uan}{y\'ing}{cheerful}{to receive}{welcome}\zhtsX%
  \tblc Please come in. 
      & \zhtss{請}{qi\v{n}g}{to request} \zhtss{進}{j\`in}{to enter}\zhtsP%
  \tblh Do you mind?
      & \zhtss{你}{n\v{i}}{you}
        \zhtsd{介}{意}{ji\`e}{y\`i}{introduce}{meaning}{to take offense}
        \zhtss{嗎}{ma}{(?)}\zhtsQ%
  \tblc &
\end{tblss}

\begin{tbls}{Anti-civilities}{sen:anti-civilities}
  \tblx Get out of here!
      & \zhtss{滾}{g\v{u}n}{to roll}%
        \zhtsd{出}{去}{ch\=u}{q\`u}{to exit}{to go}{to get out}\zhtsX%
        \citetbl{\citerp{ugly_duckling}{22}}%
  \tblh You had better kick him out as soon as possible.
      & \zhtss{你}{n\v{i}}{you}%
        \zhtsd{還}{是}{h\'ai}{sh\`i}{still}{that}{had better}%
        \zhtsd{儘}{早}{ji\v{n}}{z\v{a}o}{to the utmost}{early}{as sooon as possible}%
        \zhtssp{把}{b\v{a}}{} \zhtss{他}{t\=a}{he}%
        \zhtsd{趕}{走}{g\v{a}n}{z\v{o}u}{to expel}{to walk}{kick out}%
        \zhtssp{吧}{ba}{}\zhtsP%
        \citetbl{\citerp{ugly_duckling}{6}}
\end{tbls}


%=======================================
\section{Communications}
%=======================================
\begin{tblvv}{Common verbs}{voc:comverbs}
  \tblx to say        & \zhtss {說}{sh\=uo}{to say}
  \tblc to talk       & \zhtsds{講}{話}{ji\v{a}ng}{h\`ua}{to speak}{talks}{to talk}
  \tblh to talk about & \zhtsds{談}{論}{t\'an}{l\`un}{to talk}{to discuss}{to talk about}
  \tblc to discuss    & \zhtsds{討}{論}{t\v{a}o}{l\`un}{to demand}{to discuss}{to discuss}
  \tblh to chat       & \zhtsds{聊}{天}{li\'ao}{ti\=an}{merely}{sky}{to chat}
  \tblc to chat       & \zhtsds{談}{天}{t\'an}{ti\=an}{to talk}{sky}{to chat}
  \tblh to quarrel    & \zhtsds{吵}{架}{ch\v{a}o}{ji\`a}{to quarrel}{a rack}{to quarrel}
  \tblc to shout      & \zhtsds{喊}{叫}{h\v{a}n}{ji\`ao}{to shout}{to call}{to shout}
  \tblh indiscriminate talk & \zhtsds{亂}{話}{l\`uan}{h\`ua}{disorder}{talk}{indiscriminate talk}
  \tblc garbage talk  & \zhtsds{廢}{話}{f\`ei}{h\`ua}{abolish}{talk}{garbage talk}
  \tblh listen        & \zhtss{聽}{ti\=ng}{listen}
  \tblc listen to     & \zhtsds{聽}{著}{ti\=ng}{zhe}{listen}{(adverbial particle)}{listen to}
  \tblh listening comprehension & \zhtsds{聽}{力}{ti\=ng}{l\`i}{listen}{strength}{listening comprehension}
  \tblc sing          & \zhtsds{唱}{歌}{ch\`ang}{g\=e}{to sing}{song}{to sing}
  \tblh response      & \zhtsd{反}{應}{f\v{a}n}{yi\`ng}{contrary}{to answer}{response}
\end{tblvv}

\begin{tblv}{Expressions}{voc:expressions}
  \tblx to chat & 
    \zhtsds{談}{天}{t\'an}{ti\=an}{to talk}{sky}{to chat}
    \zhtss{說}{sh\=uo}{to say}\zhtss{笑}{xi\`ao}{to laugh}
    \citetbl{\citerp{lovemind_rain}{25}}
  \tblh to chat about everthing & 
    \zhtsds{談}{天}{t\'an}{ti\=an}{to talk}{sky}{to chat}
    \zhtss{說}{sh\=uo}{to say}\zhtss{地}{d\`i}{earth}
  \tblh everyone talking at once &
    \zhtss{七}{q\=i}{seven} \zhtss{嘴}{z\v{u}i}{mouth}
    \zhtss{八}{b\=a}{eight} \zhtss{舌}{sh\'e}{tongue} \citex{ugly_duckling}{4}
  \tblh to repeatedly urge &
    \zhtsds{叮}{嚀}{di\=ng}{ni\'ng}{to sting}{to instruct}{to repeatedly urge}
    \citex{ugly_duckling}{4}
  \tblh to say to oneself &
    \zhtss{自}{\`zi}{self}\zhtss{言}{y\'an}{speech}%
    \zhtss{自}{\`zi}{self}\zhtss{語}{y\v{u}}{language}%
    \citetbl{\citerp{baobao_bear}{21}}
  \tblh to self-ask self-answer &
    \zhtsd{自}{問}{\`zi}{w\`en}{self}{ask}{to ask oneself}%
    \zhtss{自}{\`zi}{self}\zhtss{答}{d\'a}{to answer}
  \tblh to cover the mouth with the hand &
    \zhtss{摀}{w\v{u}}{to cover}\zhtss{住}{zh\`u}{to cease}%
    \zhtsd{嘴}{巴}{z\v{u}i}{b\=a}{mouth}{to expect}{mouth}%
    \citetbl{\citerp{lovemind_octopus}{19}}
\end{tblv}

\begin{tbls}[3\tw/16]{Can you speak?}{sen:canspeak}
  \tblx Can you speak English?
      & \zhtss{你}{n\v{i}}{you}\zhtss{會}{h\`ui}{able}\zhtss{講}{ji\v{a}ng}{speak}%
        \zhtsd{英}{語}{yi\=ng}{y\v{u}}{outstanding}{language}{English}%
        \zhtss{嗎}{ma}{(?)}\zhtsQ%
  \tblh Can you speak Chinese?
      & \zhtss{你}{n\v{i}}{you}\zhtss{會}{h\`ui}{able}\zhtss{講}{ji\v{a}ng}{speak}%
        \zhtsd{國}{語}{g\'uo}{y\v{u}}{national}{language}{Chinese}%
        \zhtss{嗎}{ma}{(?)}\zhtsQ%
  \tblh I can.
      & \zhtss{會}{h\`ui}{able}\zhtsP%
  \tblh I cannot.
      & \zhtss{不}{b\`u}{not}\zhtss{會}{h\`ui}{able}\zhtsP%
  \tblh A little.
      & \zhtss{一}{\=yi}{one}\zhtss{點}{di\v{a}n}{drop}\zhtss{點}{di\v{a}n}{drop}\zhtsP%
  \tblh Just a little.
      & \zhtss{一}{y\=i}{one}\zhtss{點}{di\v{a}n}{drop}\zhtss{點}{di\v{a}n}{drop}%
        \zhtsd{而}{已}{e\'r}{y\v{i}}{and}{already}{that is all}\zhtsP%
  \tblh I can speak Chinese.
      & \zhtss{我}{w\v{o}}{I}\zhtss{會}{h\`ui}{able}\zhtss{講}{ji\v{a}ng}{speak}%
        \zhtsd{國}{語}{g\'uo}{y\v{u}}{national}{language}{Chinese}\zhtsP%
  \tblh I can speak Chinese a little.
      & \zhtss{我}{w\v{o}}{I}\zhtss{會}{h\`ui}{able}\zhtss{講}{ji\v{a}ng}{speak}%
        \zhtsd{國}{語}{g\'uo}{y\v{u}}{national}{language}{Chinese}%
        \zhtss{一}{y\=i}{one} \zhtss{點}{di\v{a}n}{drop} \zhtss{點}{di\v{a}n}{drop}\zhtsP%
  \tblh I cannot speak Chinese.
      & \zhtss{我}{w\v{o}}{I}\zhtss{不}{b\`u}{not}\zhtss{會}{h\`ui}{able}%
        \zhtss{講}{ji\v{a}ng}{speak}%
        \zhtsd{國}{語}{g\'uo}{y\v{u}}{national}{language}{Chinese}\zhtsP
\end{tbls}







