%============================================================================
% LaTeX File
% Daniel J. Greenhoe
%============================================================================

%=======================================
\section {You are my flower} % (\zht{你是我的花朵})}
%=======================================
\rule{\textwidth}{1mm}
  \zhtss{你}{n\v{i}}{you}\zhtss{是}{\`shi}{are}\zhtsd{我}{的}{w\v{o}}{de}{I}{'s}{my}%
  \zhtsd{花}{朵}{h\=ua}{d\v{u}o}{flower}{(quantifier)}{flower}
  \hspace{4ex} by 
  \zhtss{伍}{w\v{u}}{army}\zhtss{佰}{b\v{a}i}{hundred}
  \\\rule{\textwidth}{1mm}
%---------------------------------------

\begin{tbls}{You are my Flower}{song:huadou}
  \tblx Take heed to your footsteps and see everything clearly.
      & \zhtsd{留}{心}{li\'u}{xi\=n}{to remain}{heart}{to take heed}%
        \zhtsd{腳}{步}{ji\v{a}o}{b\`u}{foot}{step}{footstep}%
        \zhtss{看}{k\`an}{to see}
        \zhtss{得}{d\'e}{to the extent that}%
        \zhtsd{清}{楚}{qi\=ng}{ch\v{u}}{pure}{clear}{clear}%
  \tblh There are many dubious voices
      & \zhtss{有}{y\v{o}u}{have}%
        \zhtss{很}{h\v{e}n}{very}%
        \zhtss{多}{d\=uo}{many}%
        \zhtsd{可}{疑}{k\v{e}}{\'yi}{to approve}{doubt}{suspicious}\zhtss{的}{de}{'s}%
        \zhtsd{因}{素}{\=yin}{s\`u}{cause}{plain}{a cause}
  \tblh calling out to you, making you confused.
      & \zhtsd{呼}{喊}{h\=u}{h\v{a}n}{to exhale}{to shout}{to call out}%
        \zhtss{你}{\v{n}i}{you}%
        \zhtss{讓}{r\`{a}ng}{to cause}%
        \zhtss{你}{\v{n}i}{you}%
        \zhtsd{糊}{塗}{h\'u}{t\'u}{glue}{to smear}{confused}
  \tblh Test the extent of my love.
      & \zhtsd{考}{驗}{k\v{a}o}{y\`an}{to give an exam}{to examine}{to test}%
        \zhtsd{我}{愛}{w\v{o}}{\`ai}{I}{love}{my love}\zhtss{的}{de}{'s}%
        \zhtsd{程}{度}{ch\'eng}{d\`u}{journey}{degree}{extent}
  \tblh The elevator of a tall building in a dark neon-lit place
      & \zhtss{高}{g\=ao}{tall}\zhtss{樓}{l\'ou}{building}%
        \zhtsd{電}{梯}{di\`an}{t\=i}{electric}{ladder}{elevator}%
        \zhtsd{霓}{虹}{\'ni}{h\'ong}{secondary rainbow}{rainbow}{neon light}%
        \zhtss{深}{sh\=en}{profound}\zhtss{處}{ch\`u}{a place}
  \tblh dances in a crowd of people awkwardly,
      & \zhtsd{舞}{動}{w\v{u}}{d\`ong}{to dance}{to move}{to dance}%
        \zhtss{在}{z\`ai}{at}%
        \zhtsd{人}{群}{r\'en}{q\'un}{person}{crowd}{crowd}%
        \zhtss{很}{h\v{e}n}{very}%
        \zhtsd{拘}{束}{j\=u}{sh\`u}{to detain}{to bind}{awkward}
  \tblh pulling me so that I don't float away.
      & \zhtss{拉}{l\=a}{to pull}\zhtssp{著}{zhe}{(adv. particle)}%
        \zhtss{我}{w\v{o}}{I}%
        \zhtss{不}{b\'u}{not}\zhtss{要}{y\`ao}{want}%
        \zhtsd{飄}{走}{pi\=ao}{z\v{o}u}{to float}{to walk}{to float away}
  \tblh I am standing behind you, shielding you from the wind.
      & \zhtss{我}{w\v{o}}{I}%
        \zhtss{在}{z\`ai}{at}%
        \zhtss{你}{\v{n}i}{you}%
        \zhtsd{背}{後}{b\`ei}{h\`ou}{back}{behind}{behind}%
        \zhtss{擋}{d\v{a}ng}{to block}\zhtss{風}{f\=eng}{wind}%
  \tblh Oh you are my flower!
      & \zhtss{喔}{o}{(a sound)}%
        \zhtss{你}{n\v{i}}{you}\zhtss{是}{\`shi}{are}%
        \zhtsd{我}{的}{w\v{o}}{de}{I}{'s}{my}%
        \zhtsd{花}{朵}{h\=ua}{d\v{u}o}{flower}{(quantifier)}{flower}
  \tblh I want to have you rooted in my heart of hearts.
      & \zhtss{我}{w\v{o}}{I}\zhtss{要}{y\`ao}{want}%
        \zhtsd{擁}{有}{y\v{o}ng}{y\v{o}u}{to embrace}{to have}{to have}%
        \zhtss{你}{\v{n}i}{you}%
        \zhtss{插}{ch\=a}{to stick into}%
        \zhtss{在}{z\`ai}{at}%
        \zhtss{我}{w\v{o}}{my}%
        \zhtsd{心}{窩}{xi\=n}{w\=o}{heart}{nest}{heart of hearts}
  \tblh I want to protect you, giving you a smooth road free of obstacles.
      & \zhtss{我}{w\v{o}}{I}\zhtss{要}{y\`ao}{want}%
        \zhtsd{保}{護}{b\v{a}o}{h\`u}{to keep}{to defend}{to protect}%
        \zhtss{你}{\v{n}i}{you}%
        \zhtsd{一}{路}{\'yi}{l\`u}{one}{path}{all the way}%
        \zhtss{都}{d\=ou}{all}%
        \zhtsd{暢}{通}{ch\`ang}{t\=ong}{unobstructed}{through}{freely flowing}
  \tblh Even if the path by your side is littered with stones,
      & \zhtsd{就}{算}{ji\`u}{s\`uan}{(aux. verb)}{to calculate}{even if}%
        \zhtss{你}{\v{n}i}{you}%
        \zhtsd{身}{邊}{sh\=en}{bi\=an}{body}{side}{side}%
        \zhtss{很}{h\v{e}n}{very}\zhtss{多}{d\=uo}{many}%
        \zhtss{小}{xi\v{a}o}{small}%
        \zhtsd{石}{頭}{\'shi}{t\'ou}{stone}{head}{stones}%
  \tblh I want to sleeplessly and endlessly love you.
      & \zhtss{我}{w\v{o}}{I}\zhtss{要}{y\`ao}{want}%
        \zhtss{愛}{\`ai}{love}\zhtssp{著}{zhe}{(adv. particle)}
        \zhtss{你}{\v{n}i}{you}%
        \zhtsd{不}{眠}{b\`u}{mi\'an}{not}{sleep}{sleepless}%
        \zhtss{也}{y\v{e}}{also}%
        \zhtsd{不}{休}{b\`u}{xi\=u}{not}{rest}{endlessly}
  \tblh No matter if your heart is full of so much confusion,
      & \zhtsd{不}{管}{b\`u}{gu\v{a}n}{not}{tube}{no matter}%
        \zhtsd{你}{心}{\v{n}i}{xi\=n}{you}{heart}{your heart}%
        \zhtsd{充}{滿}{ch\=ong}{m\v{a}n}{full}{full}{full}%
        \zhtsd{多}{少}{d\=uo}{sh\v{a}o}{many}{few}{so many}%
        \zhtsd{困}{惑}{k\`un}{h\`uo}{distressed}{bewildered}{bewildered}
  \tblh I am totally unable to release myself from loving you.
      & \zhtss{我}{w\v{o}}{I}%
        \zhtsd{絕}{對}{j\'ue}{d\`ui}{extremely}{correct}{absolutely}%
        \zhtsd{不}{會}{b\'u}{h\`ui}{not}{able}{unable}%
        \zhtss{對}{d\`ui}{for}%
        \zhtss{愛}{a\`i}{love}%
        \zhtsd{放}{鬆}{f\`ang}{s\=ong}{to set free}{to loosen}{to release}
  \tblh You are the source of all the happiness that I possess.
      & \zhtss{你}{\v{n}i}{you}%
        \zhtss{是}{\`shi}{are}%
        \zhtss{我}{w\v{o}}{I}%
        \zhtsd{所}{有}{s\v{u}o}{y\v{o}u}{that which}{to have}{possess}%
        \zhtsd{快}{樂}{k\`uai}{l\`e}{quick}{joyous}{happiness}%
        \zhtsd{源}{頭}{y\'uan}{t\'ou}{fountainhead}{head}{fountainhead}%
  \tblh You and I together can reach the paradise of joy.
      & \zhtss{我}{w\v{o}}{I}%
        \zhtss{和}{h\`an}{and}%
        \zhtss{你}{\v{n}i}{you}%
        \zhtss{將}{ji\=ang}{will}%
        \zhtss{會}{h\`ui}{able}%
        \zhtsd{到}{達}{d\`ao}{d\'a}{to reach}{to reach}{to arrive at}%
        \zhtss{那}{n\`a}{that}%
        \zhtsd{開}{心}{k\=ai}{xi\=n}{to open}{heart}{joyful}%
        \zhtsd{樂}{園}{l\`e}{y\'uan}{joyous}{garden}{paradise}
\end{tbls}

%mp3 file: \url{http://mp3.ge520.com/wma/004/13/11.mp3}

%\begin{zhtzye}\Large
%你是我的花朵
%\\
%\\留心腳步看得清楚
%\\有很多可疑的因素
%\\呼喊你讓你糊涂
%\\考驗我愛的程度
%\\高樓電梯霓虹深處
%\\舞動在人群很拘束
%\\拉著我不要飄走
%\\我在你背後擋風
%\\喔你是我的花朵
%\\我要擁有你插在我心窩
%\\我要保護你一路都暢通
%\\就算你身邊很多小石頭
%\\我要愛著你不眠也不休
%\\不管你心充滿多少困惑
%\\我絕對不會對愛你放鬆
%\\你是我所有快樂源頭
%\\我和你將會到達那開心樂園
%\end{zhtzye}







