%============================================================================
% LaTeX File
% Daniel J. Greenhoe
%============================================================================

%--------------------------------------
%\section {Bright Moon}  
%--------------------------------------
%蘇軾˙水調歌頭 / 鄧麗君 唱 / 駱英 編輯

\begin{tblsn}[\tw/4]{Love Tree}{song:lovetree}
    The seed of love buried in the ground.
    %{\fntLibertineR 사랑의 땅에 씨앗 아래에 묻혀}
    %{\fntFreeSerif 사랑의 땅에 씨앗 아래에 묻혀}
    %{\fntKoreanSSK 사랑의 땅에 씨앗 아래에 묻혀}
    &
    \chars{愛}{a/`i}{love}\chars{的}{de}{'s}\chard{種}{子}{zh/>ong}{z/>i}{kind of}{son}{seed}\chars{埋}{m/'ai}{buried}%
    \chars{在}{z/`ai}{in}\chars{地}{d/`i}{ground}\chard{底}{下}{d/>i}{xi/`a}{bottom}{under}{under}
    %\chars{明}{mi\'ng}{bright} \chars{月}{y\`ue}{moon} 
    %\chard{幾}{時}{j\v{i}}{sh\'i}{how many}{time}{when} \chars{有}{y\v{o}u}{have} ？
  \tblh
    Rainwater moistens and sunlight sways.
    &
    \chard{雨}{水}{y/>u}{sh/>ui}{rain}{water}{rainwater}\chard{滋}{潤}{z/-i}{r/`un}{nourish}{moisten}{moisten}%
    \chard{陽}{光}{y/'ang}{gu/-ang}{sun}{light}{sunlight}%
    \chard{揮}{灑}{h/-ui}{s/>a}{play}{sprinkle}{sways}%
  \tblh
    Little by little new green shoots emerge.
    &
    \chard{一}{點}{y/`i}{di/>an}{one}{point}{little}%
    \chard{一}{點}{y/`i}{di/>an}{one}{point}{little}%
    \chard{冒}{出}{m/`ao}{ch/-u}{take}{out}{emerge}%
    \chars{新}{x/-in}{new}\chard{綠}{芽}{l/`:u}{y/'a}{green}{bu}{shoots}%
  \tblh
    Inch by inch struggling to climb.
    &
    \chard{一}{吋}{y/'i}{c/`un}{one}{inch}{inch}%
    \chard{一}{吋}{y/'i}{c/`un}{one}{inch}{inch}%
    \chard{奮}{力}{f/`en}{l/`i}{excited}{force}{struggle}%
    \chars{向}{xi/`ang}{to}\chard{上}{爬}{sh/`ang}{p/'a}{up}{climb}{climb}
  \tblh
    Day after day the wind blows strong.
    &
    \chard{日}{復}{r/`i}{f/`u}{day}{complex}{day after}%
    \chard{一}{日}{y/`i}{r/`i}{one}{day}{day}%
    \chard{迎}{風}{y/'ing}{f/-eng}{towards}{wind}{windward}%
    \chard{長}{大}{zh/>ang}{d/`a}{grow}{big}{grows}%
  \tblh
    Branches and leaves gradually gather into an umbrella.
    &
    \chard{枝}{葉}{zh/-i}{w/`u}{branch}{leaf}{foilage}%
    \chard{漸}{漸}{ji/`an}{ji/`an}{gradually}{gradually}{gradually}%
    \chard{聚}{成}{j/`u}{ch/'eng}{gather}{into}{gather into}%
    \chars{傘}{s/>an}{umbrella}\chard{一}{把}{y/`i}{b/>a}{one}{handle}{one}
  \tblh
    Twilight overlooks the hillside at sunset.
    &
    \chard{黃}{昏}{hu/'ang}{h/-un}{yellow}{dusk}{twilight}%
    \chard{遙}{望}{y/'ao}{w/`ang}{remote}{look}{overlook}%
    \chard{山}{邊}{sh/-an}{bi/-an}{hill}{side}{hillside}%
    \chard{晚}{霞}{w/>an}{xi/'a}{late}{---}{sunset}%
  \tblh
    At nightfall see the stars blinking.
    &
    \chard{入}{夜}{r/`u}{y/`e}{enter}{night}{At nightfall}%
    \chard{探}{看}{t/`an}{k/`an}{probe}{see}{probe}%
    \chard{星}{空}{x/-ing}{k/-ong}{star}{space}{the stars}%
    \chars{眨}{zh/>a}{blink}\chars{呀}{ya}{ah}\chars{眨}{zh/>a}{blink}%
  \tblh
    It matters not that weeds are dense and many,
    &
    \chard{不}{管}{b/`u}{gu/>an}{no}{matter}{regardless}%
    \chard{野}{草}{y/>e}{c/>ao}{wild}{grass}{weeds}%
    \chard{茂}{密}{m/`ao}{m/`i}{--}{secret}{dense}%
    \chard{紛}{雜}{f/-en}{z/'a}{numerous}{mixed}{confused}
  \tblh
    seedlings resist the pressure trying to tie them down.
    &
    \chard{根}{苗}{g/-en}{mi/'ao}{roots}{seedlings}{root and shoot}%
    \chard{抗}{壓}{k/`ang}{y/-a}{resist}{press}{resist compession}%
    \chard{努}{力}{n/>u}{l/`i}{exert}{force}{strive}%
    \chard{向}{下}{xi/`ang}{xi/`a}{to}{down}{down}\chars{紮}{z/-a}{tie}%
  \tblh
    Sometimes clouds obscure the tower of the soul.
    &
    \chard{有}{時}{y/>ou}{sh/'i}{have}{time}{sometimes}%
    \chard{雲}{霧}{y/'un}{w/`u}{cloud}{fog}{clouds}%
    \chard{遮}{蔽}{zh/-i}{b/`i}{cover}{shield}{obscure}%
    \chard{靈}{魂}{l/'ing}{h/'un}{spirit}{soul}{soul}%
    \chars{之}{zh/-i}{'s}%
    \chars{塔}{t/>a}{tower}%
  \tblh
    Sometimes it's hard not to be lonely and afraid.
    &
    \chard{有}{時}{y/>ou}{sh/'i}{have}{time}{sometimes}%
    \chard{難}{免}{n/'an}{mi/>an}{difficult}{avoid}{hard to avoid}%
    \chard{孤}{獨}{g/-u}{d/'u}{lone}{alone}{lonely}%
    \chard{害}{怕}{h/`ai}{p/`a}{harm}{fear}{afraid}%
  \tblh
    Only affection can break through the clouds
    &
    \chard{唯}{有}{w/'ei}{y/>ou}{only}{have}{only}%
    \chard{深}{情}{sh/-en}{q/'ing}{deep}{feelings}{affection}%
    \chard{能}{夠}{n/'eng}{g/`ou}{can}{enough}{can}%
    \chard{穿}{透}{chu/-an}{t/`ou}{pierce}{penetrate}{penetrate}%
    \chard{雲}{霧}{y/'un}{w/`u}{cloud}{fog}{clouds}%
  \tblh
    and bring heaven's limitless blessing.
    &
    \chard{帶}{來}{d/`ai}{l/'ai}{take}{come}{bring}%
    \chard{天}{堂}{ti/-an}{t/'ang}{sky}{hall}{heaven}%
    \chars{裡}{l/>i}{in}%
    \chars{的}{de}{'s}%
    \chard{無}{限}{w/'u}{xi/`an}{without}{limit}{limitless}%
    \chard{祝}{福}{zh/`u}{f/'u}{wish}{blessing}{blessing}%
  \tblh
    The branches of the love tree
    &
    \chard{愛}{情}{a/`i}{q/'ing}{love}{feelings}{love}%
    \chars{樹}{sh/`u}{tree}\chars{的}{de}{'s}\chard{枝}{枒}{zh/-i}{y/-a}{branch}{---}{branches}
  \tblh
    in the end bloom into a beautiful flower.
    &
    \chars{終}{zh/-ong}{in the end}%
    \chars{就}{ji/`u}{to}
    \chard{開}{出}{k/-ai}{ch/-u}{open}{out}{open up}%
    \chard{一}{朵}{y/`i}{d/>uo}{one}{---}{a}%
    \chard{美}{麗}{m/>ei}{l/`i}{beauty}{beautiful}{beautiful}%
    \chars{的}{de}{'s}%
    \chars{花}{h/-ua}{flower}%
  \tblh
\end{tblsn}
\vfill
\begin{tabular}{>{\color{green}}cl}
    \imark&\url{https://www.youtube.com/watch?v=QK2u4Ev9lwk}
  \\\imark&\url{https://www.youtube.com/watch?v=6dwAPQCBHds}
  \\\imark&\url{https://www.youtube.com/watch?v=w82Zpst6swQ}
\end{tabular}






