%============================================================================
% LaTeX File
% Daniel J. Greenhoe
%============================================================================

%--------------------------------------
\section {Bright Moon}  
\label{sec:brightmoon}
%--------------------------------------
%蘇軾˙水調歌頭 / 鄧麗君 唱 / 駱英 編輯

\begin{tbls}[\tw/4]{Bright Moon}{song:brightmoon}
    ``When did we first have the bright moon?",
    &
    \zhtss{明}{mi\'ng}{bright} \zhtss{月}{y\`ue}{moon} 
    \zhtsd{幾}{時}{j\v{i}}{sh\'i}{how many}{time}{when} \zhtss{有}{y\v{o}u}{have}\zhtsQ
  \tblh
    holding a wine glass, I asked the blue sky.
    &
    \zhtsd{把}{酒}{b\v{a}}{ji\v{u}}{grip}{alcohol}{hold a wine glass} 
    \zhtss{問}{w\`en}{ask} \zhtsd{青}{天}{qi\=ng}{ti\=an}{blue}{sky}{blue sky}\zhtsP
  \tblh
    I wonder, in Heaven's imperial palace,
    &
    \zhtsd{不}{知}{b\`u}{zh\=i}{not}{know}{wonder if}
    \zhtsd{天}{上}{ti\=an}{sh\`ang}{sky}{above}{heaven}
    \zhtsd{宮}{闕}{g\=ong}{q\=ue}{palace}{palace}{imperial palace}\zhtsC
  \tblh
    what year it is today.
    &
    \zhtsd{今}{夕}{ji\=n}{x\`i}{present}{sunset}{today} \zhtss{是}{sh\`i}{is} 
    \zhtsd{何}{年}{h\'e}{ni\'an}{what}{year}{what year}\zhtsP
  \tblh
    I desire to ride the wind to go back,
    &
   \zhtss{我}{w\v{o}}{I} \zhtss{欲}{y\`u}{desire} \zhtss{乘}{ch\'eng}{ride}
   \zhtss{風}{f\=eng}{wind} \zhtsd{歸}{去}{g\=ui}{q\`u}{to go back to}{go}{go back to}\zhtsC
  \tblh
    but dread the magnificent jade palace,
    &
    \zhtss{又}{y\`ou}{but} \zhtss{恐}{k\v{o}ng}{dread} 
    \tcom{\zhtss{瓊}{qi\'ong}{jade} \zhtss{樓}{l\'ou}{building} \zhtsd{玉}{宇}{y\`u}{y\v{u}}{jade}{roof}{splendid hall}\zhtsC}
         {magnificent jade palace}
  \tblh
    that such a high place is too cold.
    &
    \zhtss{高}{g\=ao}{high} \zhtss{處}{ch\`u}{place} 
    \zhtsd{不}{勝}{b\`u}{sh\=eng}{not}{competent}{too} \zhtss{寒}{h\'an}{cold}\zhtsP
  \tblh
    Rise and dance until making a pure shadow,
    &
    \zhtsd{起}{舞}{q\v{i}}{w\v{u}}{rise}{dance}{rise and dance} 
    \zhtss{弄}{n\`ong}{make}
    \zhtsd{清}{影}{qi\=ng}{yi\v{n}g}{pure}{shadow}{pure shadow}\zhtsC
  \tblh
    in what seems like the world of man!
    &
    \zhtsd{何}{似}{h\'e}{s\`i}{what}{similar}{similiar to} \zhtss{在}{z\`ai}{at}
    \zhtsd{人}{間}{r\'en}{ji\=an}{man}{between}{world of man}\zhtsX
  \tblh
    Arching over the bright red pavillion,
    &
    \zhtss{轉}{zh\v{u}an}{revolve} \zhtss{朱}{zh\=u}{bright red} \zhtss{閣}{g\'e}{pavillion}\zhtsC
  \tblh
    the moonlight hangs down on the beautiful door,
    &
    \zhtss{低}{d\=i}{hang down} \zhtss{綺}{q\v{i}}{beautiful} \zhtss{戶}{h\`u}{door}\zhtsC
  \tblh
    shining on the sleepless.
    &
    \zhtss{照}{zh\`ao}{illuminate} \zhtss{無}{w\'u}{without} \zhtss{眠}{mi\'an}{sleep}\zhtsP
  \tblh
    The moon should not have hate,
    &
    \zhtss{不}{b\`u}{not} \zhtsd{應}{有}{yi\=ng}{y\v{o}u}{should}{have}{deserved} 
    \zhtss{恨}{h\`en}{hate}\zhtsC
  \tblh
    so why when it is time to leave, is the moon so full and round?
    &
    \zhtsd{何}{事}{h\'e}{sh\`i}{what}{affairs}{why}
    \zhtsd{長}{向}{ch\'ang}{xi\`ang}{long}{hitherto}{always}
    \zhtsd{別}{時}{b\'ie}{sh\'i}{leave}{time}{time to leave} 
    \zhtss{圓}{y\'uan}{round}\zhtsQ
  \tblh
    People have sorrow, joy, separation, togetherness;
    &
    \zhtss{人}{r\'en}{people} \zhtss{有}{y\v{o}u}{have}
    \zhtss{悲}{b\=ei}{sadness} \zhtss{歡}{h\=uan}{cheerful} 
    \zhtss{離}{l\'i}{apart} \zhtss{合}{h\'e}{together}\zhtsC
  \tblh
    the moon has times of being overcast, clear, full and round, diminished,
    &
    \zhtss{月}{y\`ue}{moon} \zhtss{有}{y\v{o}u}{have} \zhtss{陰}{y\=in}{overcast}
    \zhtss{晴}{qi\'ng}{clear} \zhtss{圓}{y\'uan}{round} \zhtss{缺}{q\=ue}{deficient}\zhtsC
  \tblh
    since ancient times, these things have been difficult to reconcile.
    &
    \zhtss{此}{c\v{i}}{these} \zhtss{事}{sh\`i}{affairs} \zhtss{古}{g\v{u}}{ancient} 
    \zhtss{難}{n\'an}{difficult} \zhtss{全}{q\'uan}{completely}\zhtsP
  \tblh
    If only people can remain together for a long time,
    &
    \zhtsd{但}{願}{d\`an}{y\`uan}{but}{hope}{if only}
    \zhtss{人}{r\'en}{people}
    \zhtsd{長}{久}{ch\'ang}{ji\v{u}}{long}{time}{long time}\zhtsC
  \tblh
    or travel a thousand miles to share life with a lovely lady.
    &
    \zhtsd{千}{里}{qi\=an}{l\v{i}}{thousand}{li}{long distance}
    \zhtss{共}{g\`ong}{share} 
    \zhtsd{嬋}{娟}{ch\'an}{j\=uan}{graceful}{beautiful}{lovely}\zhtsP
\end{tbls}

midi file available at:
\url{http://www.hamienet.com/midi15005.html}







